OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-03-05 paper/paper.Rmd
Describe data processing

To evaluate reference-based clustering with independent databases, we fit each
dataset to reference OTUs generated by _de novo_ clustering the Greengenes
database (v13_8_99), Silva non-redundant database (v132), or the Ribosomal
Database Project (RDP; v16). **TODO: cite database papers?** In open reference
mode, fitting the datasets to reference OTUs with OptiFit produced OTUs of
similar quality as clustering the datasets _de novo_ with OptiClust (**TODO:
diff. in medians**). This held true for all datasets and reference databases.
OptiFit produced higher quality OTUs than vsearch when open reference clustering
against the greengenes database, with median MCC scores of **A** and **B%**
(respectively). However, open reference vsearch and OptiClust respectively ran
**X%** and **Y%** faster than OptiFit in open reference mode.

@@

>>>>

To evaluate reference-based clustering with independent databases, we fit each
dataset to reference OTUs generated by _de novo_ clustering the Greengenes
database, Silva non-redundant database, or the Ribosomal
Database Project (RDP). **TODO: cite database papers?** In open reference
mode, fitting the datasets to reference OTUs with OptiFit produced OTUs of
similar quality as clustering the datasets _de novo_ with OptiClust (**TODO:
diff. in medians**). This held true for all datasets and reference databases.
OptiFit produced higher quality OTUs than vsearch when open reference clustering
against the greengenes database, with median MCC scores of **A** and **B%**
(respectively). However, open reference vsearch and OptiClust respectively ran
**X%** and **Y%** faster than OptiFit in open reference mode.

@@

We downloaded 16S rRNA gene amplicon sequences from four published datasets
isolated from soil [@johnston_metagenomics_2016], marine
[@henson_artificial_2016], mouse gut [@schloss_stabilization_2012], and human
gut [@baxter_microbiota-based_2016] samples.
Raw sequences were processed using mothur according to the Schloss Lab MiSeq SOP
as described in the mothur wiki and accompanying study by Kozich _et al._
[@schloss_miseq_nodate; @kozich_development_2013].
These steps included trimming and filtering for quality, aligning to the Silva
reference alignment [**CITE**], discarding sequences that aligned outside the V4
region, removing chimeric reads with UCHIME, and calculating distances between
all pairs of sequences within each dataset prior to clustering.
To generate reference OTUs from independent databases, we downloaded sequences
from the Greengenes database (v13_8_99) [**CITE**], Silva non-redundant database
(v132) [**CITE**], and the Ribosomal Database Project (RDP; v16) [**CITE**].
These sequences were processed using the same steps outlined above followed by
clustering sequences into _de novo_ OTUs with OptiClust.
Processed reads from each of the four datasets were clustered with OptiFit to
the reference OTUs generated form each of the three datasets.
When reference clustering with vsearch, processed datasets were fit directly
to the unprocessed greengenes reference alignment, since this is how vsearch
is typically used by the QIIME2 software [@bolyen_reproducible_2019].
